---
title: "HW 3"
editor: visual
editor_options: 
  chunk_output_type: console
---

::: callout-important
This homework is due Thursday, Oct 26 at 5:00 pm ET.
:::

# Getting Started

-   Go to the [sta113-f23](https://github.com/sta113-f23) organization on GitHub. Click on the repo with the prefix `hw-3`. It contains the starter documents you need to complete the homework assignment.
-   Click on the green **CODE** button, select **Use SSH** (this might already be selected by default, and if it is, you'll see the text **Clone with SSH**). Click on the clipboard icon to copy the repo URL.
-   In RStudio, go to *File* ➛ *New Project* ➛*Version Control* ➛ *Git*.
-   Copy and paste the URL of your assignment repo into the dialog box *Repository URL*. Again, please make sure to have *SSH* highlighted under *Clone* when you copy the address.
-   Click *Create Project*, and the files from your GitHub repo will be displayed in the *Files* pane in RStudio.
-   Click *hw-3.qmd* to open the template Quarto file. This is where you will write up your code and narrative for the lab.

## Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
```

## Guidelines + tips {#sec-guidelines-tips}

AYour plots should include an informative title, axes should be labeled, and careful consideration should be given to aesthetic choices.

Remember that continuing to develop a sound workflow for reproducible data analysis is important as you complete this homework and other assignments in this course. There will be periodic reminders in this assignment to remind you to knit, commit, and push your changes to GithHub. You should have **at least 3 commits with meaningful commit messages** by the end of the assignment.

::: callout-note
Note: Do not let R output answer the question for you unless the question specifically asks for just a plot. For example, if the question asks for the number of columns in the data set, please type out the number of columns. You are subject to lose points if you do not.
:::

## Workflow + formatting

Make sure to

-   Update author name on your document.
-   Label all code chunks informatively and concisely.
-   Follow the Tidyverse code style guidelines.
-   Make at least 3 commits.
-   Resize figures where needed, avoid tiny or huge plots.
-   Use informative title and axis labels.
-   Turn in an organized, well formatted document.

# Exercises

## Credit card balances

The data for Exercises 1-\_\_\_ is on credit card balances. The variables of interest from the dataset, called `credit`, are as follows:

The dataset is in the `data` folder of your repository, and it's called `credit.csv`. It contains the following variables:

-   `balance`: Credit card balance in \$
-   `income`: Income in \$1,000
-   `student`: Whether the individual is a student (`Yes`) or not (`No`)
-   `married`: Whether the individual is a married (`Yes`) or not (`No`)
-   `limit`: Credit limit

### Exercise 1

Load the data from a relative path.

```{r}
credit <- read_csv("data/credit.csv")
```

### Exercise 2

Make a scatterplot of `balance` vs. `income` and overlay the line of best fit, without the uncertainty band around it.

```{r}
ggplot(credit, aes(x = income, y = balance)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE)
```

### Exercise 3

Fit the regression model for predicting balance from income, display a tidy output of the model, and interpret the slope and the intercept.

```{r}
credit_fit <- linear_reg() |>
  set_engine("lm") |>
  fit(balance ~ income, data = credit)

tidy(credit_fit)
```

### Exercise 4

Plot the residuals

```{r}
credit_aug <- augment(credit_fit, new_data = credit)

ggplot(credit_aug, aes(x = .pred, y = .resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE)
```

### Exercise 5

```{r}
credit |>
  mutate(utilization = balance / limit) |>
  ggplot(aes(x = income, y = utilization, color = student, shape = student)) +
  geom_point(alpha = 0.5, show.legend = FALSE) +
  geom_smooth(method = "lm", se = FALSE, show.legend = FALSE) +
  facet_grid(
    student~married, 
    labeller = labeller(.rows = label_both, .cols = label_both)
  ) +
  scale_x_continuous(labels = label_dollar(scale = 1, suffix = "K")) +
  scale_y_continuous(labels = label_percent()) +
  scale_color_manual(values = c("steelblue", "brown3")) +
  labs(
    x = "Income",
    y = "Credit utilization",
    color = "Student",
    shape = "Student"
  ) +
  theme_bw()
```

### Exercise 6

Based on the visualization you created in Exercise 5, how, if at all, are the relationships between income and credit utilization different than the relationships between income and credit balance for individuals with various student and marriage status.

*For non-students (married and unmarried) the relationship is still positive, but the slope is not as steep as the relationship between income and balance. For students who are married the relationship is basically nonexistent. For students who are not married, the relationship is negative*

### Exercise 7



::: render-commit-push
Now is a good time to render, commit, and push. Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding. You can use a commit message like "Finished Part 1".
:::

## Miscellaneous

### Exercise 8

**Read the docs:**



::: render-commit-push
You know the drill: render, commit, and push!
:::

### Exercise 9

::: render-commit-push
Render, commit, and push one last time.

Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::

# Wrap up

## Review

Before you call it "done",

-   make sure you've answered all questions (or, at least, have not skipped any accidentally) and
-   review the @sec-guidelines-tips section and make any edits/corrections needed.

You can change answers to exercises you've "completed" and commit and push again. We will only grade the final version of your assignment. You can commit and push as many times as you like until the deadline.

## Submission

You do not have to do anything special to "submit" your assignment. We will close the repos to further pushes at the deadline, and will take your work as of that time point as your submission.

If you need to submit late for any reason, contact the professor.

## Grading

-   Exercise 1-9: 10 points each
-   Workflow + formatting: 10 points
-   Total: 100 points
