---
title: "Modelling fish weights with multiple predictors"
categories: 
  - Application exercise
editor: visual
editor_options: 
  chunk_output_type: console
format: 
  html:
    fig-width: 6
    fig-asp: 0.618
---

For this application exercise, we will continue to work with data on fish. The dataset we will use, called `fish`, is on two common fish species in fish market sales. We're going to investigate the relationship between the weights and heights of fish, and later take into consider species as well.

```{r}
#| label: load-packages-data
#| message: false

library(tidyverse)
library(tidymodels)

fish <- read_csv("data/fish.csv")
```

The data dictionary is below:

| **variable**      | **description**        |
|-------------------|------------------------|
| `species`         | Species name of fish   |
| `weight`          | Weight, in grams       |
| `length_vertical` | Vertical length, in cm |
| `length_diagonal` | Diagonal length, in cm |
| `length_cross`    | Cross length, in cm    |
| `height`          | Height, in cm          |
| `width`           | Diagonal width, in cm  |

# Interpreting multiple regression models

In the previous application exercise you saw that the model predicting weight from height and species was a better fit. In this section we will interpret the coefficients of this model.

```{r}
#| label: fit-whs

fish_whs_fit <- linear_reg() |>
  fit(weight ~ height + species, data = fish)

tidy(fish_whs_fit)
```

-   What does each row in the model output represent?

*Add response here.*

-   Interpret the intercept and the slopes.

*Add response here.*

-   Write the model.

*Add response here.*

# Additive vs. interaction models

-   Run the two code chunks below and create two separate plots. How are the two plots different than each other? Which plot does the model we fit above represent?

```{r}
#| label: model-a

ggplot(fish, aes(x = height, y = weight, color = species)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Plot A - Interaction model")
```

```{r}
#| label: model-b

fish_whs_aug <- augment(fish_whs_fit, new_data = fish)

ggplot(
  fish_whs_aug, 
  aes(x = height, y = weight, color = species)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = .pred), method = "lm", se = FALSE) +
  labs(title = "Plot B - Additive model")
```

-   Look back at Plot B. What assumption does the additive model make about the slopes between flipper length and body mass for each of the three islands?

*Add response here.*

# Choosing models

Rule of thumb: **Occam's Razor** - Don't overcomplicate the situation! We prefer the *simplest* best model.

-   Choose a model using this principle.

```{r}
#| label: model-selection

# add code here
```

*Add response here.*

-   What is R-squared? What is adjusted R-squared?

*Add response here.*
