---
title: "Donors and yawners"
categories: 
  - Application exercise
  - Suggested answers
editor: visual
editor_options: 
  chunk_output_type: console
---

In this application exercise, we'll introduce pipelines for conducting hypothesis tests with randomization.

# Goals

-   Conduct a hypothesis test for a difference in proportions

# Packages and data

We'll use the **tidyverse** and **tidymodels** packages as usual and the **openintro** package for the datasets.

```{r}
#| label: load-packages-data
#| message: false

library(tidyverse)
library(tidymodels)
library(openintro)
```

Using the `yawn` dataset in the **openintro** package, conduct a hypothesis test for evaluating whether yawning is contagious.
First, set the hypotheses.
Then, conduct a randomization test using 1000 simulations.
Visualize and calculate the p-value and use it to make a conclusion about the statistical discernability of the difference in proportions of yawners in the two groups.
Then, comment on whether you "buy" this conclusion.

## Exercise 1

Construct the null distribution with 100 resamples.
Name it `null_dist`.
How many rows does `null_dist` have?
How many columns?
What does each row and each column represent?

```{r}
#| label: null-dist

# add code here
```

*Add response here.*

## Exercise 2

Where do you expect the center of the null distribution to be?
Visualize it to confirm.

```{r}
#| label: null-dist-plot

# option 1

# add code here

# option 2

# add code here
```

## Exercise 2

Calculate the observed difference in proportions.
Name it `obs_stat`.

```{r}
#| label: obs-stat

# add code here
```

## Exercise 3

Overlay the observed statistic on the null distribution and comment on whether an observed outcome as extreme as the observed statistic, or lower, is a likely or unlikely outcome, if in fact the null hypothesis is true.

```{r}
#| label: obs-stat-plot

# option 1

# add code here

# option 2

# add code here
```

## Exercise 4

Calculate the p-value and comment on whether it provides convincing evidence that yawning is contagious.

*Add response here.*

```{r}
#| label: pvalue-donor

# option 1

# add code here

# option 2

# add code here
```

## Exercise 5

Let's get real!
Redo the test with 10,000 simulations.
Note: This can take some time to run.

```{r}
#| label: ht-more-sims

# add code here
```

## Exercise 6

Use the p-value to make a conclusion about the statistical discernability of the difference in proportions of yawners in the two groups.
Then, comment on whether you "buy" this conclusion.

*Add response here.*
